---
title: "Population Maker"
author: "Nate Lant"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
persons <- read_csv("scenarios/slc_beam/wheelchair/final_persons.csv")
households <- read_csv("scenarios/slc_beam/wheelchair/final_households.csv")
trips <- read_csv("scenarios/slc_beam/wheelchair/final_trips.csv")
```


The challenge we face is how we can accurately simulate a tiny population of 
wheelchair users compared to the enormous general population. If we take a random 5%
sample, for example, we would get a limited number of wheelchair users.

My idea is to simulate four different saturation levels of wc users and evaluate WAV usage
at each level. The first level is 100% of the wheelchair population. The second level is 100% 
wc population and a random sample of the same size from the general population. The third
level is 100% of the wc population and a random sample of the general population ten times bigger
than the wheelchair population. The fourth level is 10% sample of the entire population.

In summary, let x = full wheelchair population and y = random sample of general population and
x = y in volume and n = total population in scenario.

1. $n = x$
2. $n = x + y$
3. $n = x + 9y$
4. $n = x/10 + total_y/10$

# 1. All Wheelchair population

```{r}
# wheelchair population
wc_population <- persons %>% 
  filter(wc_var == T)

# find households
households %>% filter(household_id %in% wc_population$household_id)
```

